---
title: "Statistical Analysis and Results"
format: 
  docx:
    toc: false
    number-sections: true
    highlight-style: github
    prefer-html: true
bibliography: 
  references.bib
  grateful-refs.bib
csl: apa.csl
---

```{r} 
#| message: false
#| warning: false
#| echo: false
#| include: false


knitr::opts_chunk$set(echo = TRUE, fig.align="center")
options(knitr.kable.NA = '')

library(tidyverse)
library(kableExtra)
library(ggtext)
library(patchwork)
library(grateful)

```

# Statistical Analysis

The present analysis was not pre-registered as we had no a priori hypotheses and, given the limited sample size due to resource constraints, thus was considered exploratory. Inferential statistics were treated as highly unstable local descriptions of the relations between model assumptions and data in order to acknowledge the inherent uncertainty in drawing generalised inferences from single and small samples [@amrhein_inferential_2019]. For all analyses we opted to take an estimation-based approach typical when examining validity and reliability. Two sets of models were employed exploring the JHs recorded from CMJ trials; one to examine the agreement between the IM method using the Force Decks (i.e., gold-standard) and each of the other methods, and one to explore the test-retest reliability for each method. 

All data and code is presented in the supplementary materials (https://osf.io/e7qg5/).
```{r}
cite_packages(output = "paragraph", out.dir = ".")
```

## Agreement
We explored the agreement between the gold-standard and other methods examined using the Concordance Correlation Coefficient 

ADD - https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x 

# Results
```{r} 
#| message: false
#| warning: false
#| echo: false

targets::tar_load(tidy_disability_velocity_model)
targets::tar_load(tidy_classif_velocity_model)
targets::tar_load(tidy_disability_acceleration_model)
targets::tar_load(tidy_classif_acceleration_model)

disability_tab <- left_join(tidy_disability_velocity_model, tidy_disability_acceleration_model,
                          by = c("effect", "term")) %>%
  select(term, estimate.x, conf.low.x, conf.high.x, estimate.y, conf.low.y, conf.high.y)

classif_tab <- left_join(tidy_classif_velocity_model, tidy_classif_acceleration_model,
                            by = c("effect", "term")) %>%
  select(term, estimate.x, conf.low.x, conf.high.x, estimate.y, conf.low.y, conf.high.y)


tidy_model_sprint <- rbind(disability_tab, classif_tab) %>%
  mutate_if(is.numeric, round, 2)


tidy_model_sprint$term <- recode(tidy_model_sprint$term,
                              "(Intercept)" = "Intercept",
                              "disabilitySCI" = "$Disability_{SCI}$",
                              "distance5M10m" = "$Distance_{5-10m}$",
                              "distance10M15m" = "$Distance_{10-15m}$",
                              "distance15M20m" = "$Distance_{15-20m}$",
                              "disabilitySCI:distance5M10m" = "$Disability_{SCI}$:$Distance_{5-10m}$",
                              "disabilitySCI:distance10M15m" = "$Disability_{SCI}$:$Distance_{10-15m}$",
                              "disabilitySCI:distance15M20m" = "$Disability_{SCI}$:$Distance_{15-20m}$",
                              "sd__(Intercept)" = "$\\sigma_{Intercept}$",  
                              "sd__distance5M10m" = "$\\sigma_{distance5M10m}$",
                              "sd__distance10M15m" = "$\\sigma_{distance10M15m}$",
                              "sd__distance15M20m" = "$\\sigma_{distance15M20m}$",
                              "cor__(Intercept).distance5M10m" = "$\\rho_{Intercept:Distance_{5-10m}}$",
                              "cor__(Intercept).distance10M15m" = "$\\rho_{Intercept:Distance_{10-15m}}$",
                              "cor__(Intercept).distance15M20m" = "$\\rho_{Intercept:Distance_{15-20m}}$",
                              "cor__distance5M10m.distance10M15m" = "$\\rho_{Distance_{5-10m}:Distance_{10-15m}}$",
                              "cor__distance5M10m.distance15M20m" = "$\\rho_{Distance_{5-10m}:Distance_{15-20m}}$",
                              "cor__distance10M15m.distance15M20m" = "$\\rho_{Distance_{10-15m}:Distance_{15-20m}}$",
                              "sd__Observation" = "$\\sigma_{Residual}$",
                              
                              "classif5" = "$Classification_{5}$",
                              "classif5:distance5M10m" = "$Classification_{5}$:$Distance_{5-10m}$",
                              "classif5:distance10M15m" = "$Classification_{5}$:$Distance_{10-15m}$",
                              "classif5:distance15M20m" = "$Classification_{5}$:$Distance_{15-20m}$"
                              )

tidy_model_sprint <- tidy_model_sprint[c(1:14,16,15,17:33,35,35,36:38),]
```

```{r} 
#| message: false
#| warning: false
#| echo: false

targets::tar_load(tidy_disability_rsa_model)
targets::tar_load(tidy_classif_rsa_model)

tidy_model_rsa <- rbind(tidy_disability_rsa_model,
                         tidy_classif_rsa_model) %>%
  select(term, estimate, conf.low, conf.high)  %>%
  mutate_if(is.numeric, round, 2)


tidy_model_rsa$term <- recode(tidy_model_rsa$term,
                              "(Intercept)" = "Intercept",
                              "disabilitySCI" = "$Disability_{SCI}$",
                              "sprint_number" = "Sprint Number",
                              "distance5M10m" = "$Distance_{5-10m}$",
                              "distance10M15m" = "$Distance_{10-15m}$",
                              "distance15M20m" = "$Distance_{15-20m}$",
                              "disabilitySCI:sprint_number" = "$Disability_{SCI}$:Sprint Number",
                              "disabilitySCI:distance5M10m" = "$Disability_{SCI}$:$Distance_{5-10m}$",
                              "disabilitySCI:distance10M15m" = "$Disability_{SCI}$:$Distance_{10-15m}$",
                              "disabilitySCI:distance15M20m" = "$Disability_{SCI}$:$Distance_{15-20m}$",
                              "sprint_number:distance5M10m" = "Sprint Number:$Distance_{5-10m}$",
                              "sprint_number:distance10M15m" = "Sprint Number:$Distance_{10-15m}$",
                              "sprint_number:distance15M20m" = "Sprint Number:$Distance_{15-20m}$",
                              "disabilitySCI:sprint_number:distance5M10m" = "$Disability_{SCI}$:Sprint Number:$Distance_{5-10m}$",
                              "disabilitySCI:sprint_number:distance10M15m" = "$Disability_{SCI}$:Sprint Number:$Distance_{10-15m}$",
                              "disabilitySCI:sprint_number:distance15M20m" = "$Disability_{SCI}$:Sprint Number:$Distance_{15-20m}$",
                              
                              "sd__(Intercept)" = "$\\sigma_{Intercept}$",  
                              "sd__sprint_number" = "$\\sigma_{Sprint Number}$",
                              "sd__distance5M10m" = "$\\sigma_{distance5M10m}$",
                              "sd__distance10M15m" = "$\\sigma_{distance10M15m}$",
                              "sd__distance15M20m" = "$\\sigma_{distance15M20m}$",
                              
                              "cor__(Intercept).sprint_number" = "$\\rho_{Intercept:Sprint Number}$",
                              "cor__(Intercept).distance5M10m" = "$\\rho_{Intercept:Distance_{5-10m}}$",
                              "cor__(Intercept).distance10M15m" = "$\\rho_{Intercept:Distance_{10-15m}}$",
                              "cor__(Intercept).distance15M20m" = "$\\rho_{Intercept:Distance_{15-20m}}$",
                              
                              "cor__sprint_number.distance10M15m" = "$\\rho_{Sprint Number:Distance_{10-15m}}$",
                              "cor__sprint_number.distance5M10m" = "$\\rho_{Sprint Number:Distance_{5-10m}}$",
                              "cor__sprint_number.distance15M20m" = "$\\rho_{Sprint Number:Distance_{15-20m}}$",
                              
                              
                              "cor__distance5M10m.distance10M15m" = "$\\rho_{Distance_{5-10m}:Distance_{10-15m}}$",
                              "cor__distance5M10m.distance15M20m" = "$\\rho_{Distance_{5-10m}:Distance_{15-20m}}$",
                              "cor__distance10M15m.distance15M20m" = "$\\rho_{Distance_{10-15m}:Distance_{15-20m}}$",
                              "sd__Observation" = "$\\sigma_{Residual}$",
                              
                              "classif5" = "$Classification_{5}$",
                              "classif5:sprint_number" = "$Classification_{5}$:Sprint Number",
                              "classif5:distance5M10m" = "$Classification_{5}$:$Distance_{5-10m}$",
                              "classif5:distance10M15m" = "$Classification_{5}$:$Distance_{10-15m}$",
                              "classif5:distance15M20m" = "$Classification_{5}$:$Distance_{15-20m}$",
                              "classif5:sprint_number:distance5M10m" = "$Classification_{5}$:Sprint Number:$Distance_{5-10m}$",
                              "classif5:sprint_number:distance10M15m" = "$Classification_{5}$:Sprint Number:$Distance_{10-15m}$",
                              "classif5:sprint_number:distance15M20m" = "$Classification_{5}$:Sprint Number:$Distance_{15-20m}$",
                              
                              )

tidy_model_rsa <- tidy_model_rsa[c(1:21,22,23,25,28,24,26,29,27,30:55,57,60,56,58,61,59,62:64),]
```

```{r} 
#| message: false
#| warning: false
#| echo: false

targets::tar_load(tidy_disability_lactate_model)
targets::tar_load(tidy_classif_lactate_model)

tidy_model_lactate <- rbind(tidy_disability_lactate_model,
                         tidy_classif_lactate_model) %>%
  select(term, estimate, conf.low, conf.high)  %>%
  mutate_if(is.numeric, round, 2)


tidy_model_lactate$term <- recode(tidy_model_lactate$term,
                              "(Intercept)" = "Intercept",
                              "disabilitySCI" = "$Disability_{SCI}$",
                              "time" = "Time",
                              "disabilitySCI:time" = "$Disability_{SCI}$:Time",
                              
                              "sd__(Intercept)" = "$\\sigma_{Intercept}$",  
                              "sd__Observation" = "$\\sigma_{Residual}$",
                              
                              "class5" = "$Classification_{5}$",
                              "class5:time" = "$classication_{5}$:Time"
                              )

```

## Sprint trial outcomes

The overall grand means and credible intervals from the models for the fixed effects (i.e., without including the random effects) for both velocity and acceleration can be seen in @fig-velocity-model-plot and @fig-acceleration-model-plot, in addition to individual data, respectively for both disability and classification models. All parameters for both outcomes and both disability and classification models are also shown in @tbl-sprint-model. As might be expected, fixed effects in both models revealed that velocity increased as distance covered increased and the reverse pattern for acceleration which decreased as distance covered increased. Random effects in both models showed that variation in velocities increased with increasing distance covered, and also the random effects correlations suggested that those who were initially faster, or faster during certain sections of the sprint, were similarly typically faster at all other distances. Variance in acceleration was more similar over increasing distance covered as compared with velocity, and also the random effects correlations suggested that those who had initially higher acceleration showed greater declines in acceleration across all distances, though between adjacent distances there were more positive relationships. 

### Disability
SCI participants showed slower velocities across all distances. There was however little interaction effect between disability and distance upon velocity. SCI participants also had lower acceleration across all distances. However, there were interactions between disability and distance whereby although over the initial 0-5m distance SCI participants had lower accelerations, the difference between them and participants with other injuries decreased as distance covered increased. During the final 10-15 and 15-20m accelerations were similar between groups. 

### Classification
Both 4s and 5s showed similar velocities across all distances, as well as accelerations. There was little effect of classification upon either velocity or acceleration.  

```{r} 
#| message: false
#| warning: false
#| echo: false
#| label: tbl-sprint-model 
#| tbl-cap-location: "top"
#| tbl-cap: Model parameter estimates for both fixed and random effects for sprint trial outcomes (velocity and acceleration).


knitr::kable(
  tidy_model_sprint,
  align = c("l","c","c","c","c","c","c"),
  col.names = c("Model Term",
                "Estimate",
                "Lower 95% CI",
                "Upper 95% CI",
                "Estimate",
                "Lower 95% CI",
                "Upper 95% CI"
                )
) %>%
  footnote(general = c("CI = credible interval")
           ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE) %>%
  add_header_above(c(" ", "Velocity $\\\\(m\U00B7s^{-1})$" = 3, "Acceleration $\\\\(m\U00B7s^{-2})$" = 3), 
                   escape = FALSE) %>%
  pack_rows("Disability Model", 1,19) %>%
  pack_rows("Fixed Effects", 1,8, italic = TRUE, bold = FALSE) %>%
  pack_rows("Random Effects", 9,19, italic = TRUE, bold = FALSE) %>%
  pack_rows("Classification Model", 20,38) %>%
  pack_rows("Fixed Effects", 20,27, italic = TRUE, bold = FALSE) %>%
  pack_rows("Random Effects", 28,38, italic = TRUE, bold = FALSE) %>%
  kableExtra::kable_styling()

```



```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-velocity-model-plot 
#| fig-width: 15
#| fig-height: 7.5
#| fig-cap: Individual data (top row) and global grand means with distribution and 95% credible interval estimates from the expectation of the posterior predictive distribution (bottom row) for velocity by both disability, panels (A) and (B), and classification, panels (C) and (D).

targets::tar_load(combined_velocity_plot)

combined_velocity_plot

```
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-acceleration-model-plot 
#| fig-width: 15
#| fig-height: 7.5
#| fig-cap: Individual data (top row) and global grand means with distribution and 95% credible interval estimates from the expectation of the posterior predictive distribution (bottom row) for acceleration by both disability, panels (A) and (B), and classification, panels (C) and (D).

targets::tar_load(combined_acceleration_plot)

combined_acceleration_plot

```

## Repeated sprint trial outcomes

### Sprint times

The overall grand means and credible intervals from the models for the fixed effects (i.e., without including the random effects) for repeated sprint times can be seen in @fig-rsa-model-plot and @fig-acceleration-model-plot, in addition to individual data and participant level linear smooths, respectively for both disability and classification models. All parameters for both outcomes and both disability and classification models are also shown in @tbl-rsa-model.

On average, fixed effects in both models revealed that sprint number had little impact on time, however did interact with distance revealing greater increases in time for later sprints over increasing distances. Sprint number had little impact upon the initial 0-5m. Of course, trivially, time increased as distance covered increased. Random effects in both models showed, similarly to velocity in the sprint trials, that variation in times increased with increasing distance covered. Also the random effects correlations suggested that those who were initially faster at the beginning of a sprint, faster during certain sections of the sprint, or faster during a given sprint number, were similarly typically faster at all other distances and during all other sprint numbers. 

#### Disability
Both SCI and other disabilities showed similar performances in the repeated sprints, across all distances, and all sprint numbers. There was little effect of disability upon either repeated sprint times.

#### Classification
Both 4s and 5s showed similar performances in the repeated sprints, across all distances, and all sprint numbers. There was little effect of classification upon either repeated sprint times.

```{r} 
#| message: false
#| warning: false
#| echo: false
#| label: tbl-rsa-model 
#| tbl-cap-location: "top"
#| tbl-cap: Model parameter estimates for both fixed and random effects for repeated sprint trial times.


knitr::kable(
  tidy_model_rsa,
  align = c("l","c","c","c"),
  col.names = c("Model Term",
                "Estimate",
                "Lower 95% CI",
                "Upper 95% CI"
              )
) %>%
  footnote(general = c("CI = credible interval")
           ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE) %>%
  add_header_above(c(" ", "Time (seconds)" = 3), 
                   escape = FALSE) %>%
  pack_rows("Disability Model", 1,32) %>%
  pack_rows("Fixed Effects", 1,16, italic = TRUE, bold = FALSE) %>%
  pack_rows("Random Effects", 17,32, italic = TRUE, bold = FALSE) %>%
  pack_rows("Classification Model", 33,64) %>%
  pack_rows("Fixed Effects", 33,48, italic = TRUE, bold = FALSE) %>%
  pack_rows("Random Effects", 49,64, italic = TRUE, bold = FALSE) %>%
  kableExtra::kable_styling()

```



```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-rsa-model-plot 
#| fig-width: 15
#| fig-height: 7.5
#| fig-cap: Individual data with linear smooths by participant (top row) and global grand means with distribution and 95% credible interval estimates from the expectation of the posterior predictive distribution (bottom row) for repeated sprint times by both disability, panels (A) and (B), and classification, panels (C) and (D).

targets::tar_load(combined_rsa_plot)

combined_rsa_plot

```

### Blood lactate
The overall grand means and credible intervals from the models for the fixed effects (i.e., without including the random effects) for blood lactate can be seen in @fig-lactate-model-plot in addition to individual data, respectively for both disability and classification models. All parameters for both outcomes and both disability and classification models are also shown in @tbl-lactate-model. As might be expected, fixed effects in both models revealed that blood lactate increased as from pre- to post-repeated sprint trials (see $Time$ terms in @tbl-lactate-model). Random intercepts also showed some variation in baseline blood lactate levels. 

### Disability
There was little difference in average blood lactate levels between those with SCI or other disabilities, nor was there a clear interaction effect suggesting both groups increased in blood lactate similarly. 

### Classification
Both 4s and 5s showed similar average blood lactate levels too. However, the posterior estimates were suggestive of an interaction effect little effect of classification upon either velocity or acceleration whereby 5s tended to show a greater increase in blood lactate levels post-repeated sprint trials.  

```{r} 
#| message: false
#| warning: false
#| echo: false
#| label: tbl-lactate-model 
#| tbl-cap-location: "top"
#| tbl-cap: Model parameter estimates for both fixed and random effects for blood lactate pre- and post-repeated sprint trials.


knitr::kable(
  tidy_model_lactate,
  align = c("l","c","c","c"),
  col.names = c("Model Term",
                "Estimate",
                "Lower 95% CI",
                "Upper 95% CI"
              )
) %>%
  footnote(general = c("CI = credible interval")
           ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE) %>%
  add_header_above(c(" ", "Blood lactate ($\\\\(m\U00B7L^{-1})$)" = 3), 
                   escape = FALSE) %>%
  pack_rows("Disability Model", 1,6) %>%
  pack_rows("Fixed Effects", 1,4, italic = TRUE, bold = FALSE) %>%
  pack_rows("Random Effects", 5,6, italic = TRUE, bold = FALSE) %>%
  pack_rows("Classification Model", 7,12) %>%
  pack_rows("Fixed Effects", 7,9, italic = TRUE, bold = FALSE) %>%
  pack_rows("Random Effects", 11,12, italic = TRUE, bold = FALSE) %>%
  kableExtra::kable_styling()

```



```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-lactate-model-plot 
#| fig-width: 15
#| fig-height: 7.5
#| fig-cap: Individual data with linear smooths by participant (top row) and global grand means with distribution and 95% credible interval estimates from the expectation of the posterior predictive distribution (bottom row) for repeated sprint times by both disability, panels (A) and (B), and classification, panels (C) and (D).

targets::tar_load(combined_lactate_plot)

combined_lactate_plot

```

# References
